---
output: github_document
editor_options: 
  markdown: 
    wrap: 72
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# SamsaRaLight

The goal of SamsaRaLight is to estimate the light intercepted by a tree
and the light on the ground in a given forest stand.

It uses the ray tracing model SamsaraLight (Courbaud et al. 2003, 2015)
to compute the direct and diffuse rays within a given period of a year,
at a given latitude. Then, it simulates the ray coming from the sky
towards cells of the plot and it computes the interception of all the
rays by each consecutive tree crowns with given dimensions.

The SamsaraLight model was initially implemented within the Java
platform Capsis (<https://capsis.cirad.fr/capsis/help_en/samsaralight>).
However, for the sake of fast computing and easier use within R
paradigm, we needed to reformulate the model as an R package. We used
both `data.table` R package and Rcpp with C++ script for fast computing.

## Installation

You can install the development version of SamsaRaLight from
[GitHub](https://github.com/) with:

``` r
install.packages("devtools")
devtools::install_github("NatheoB/SamsaRaLight")
```

## Example

This is a basic example which shows you how to solve a common problem:

```{r example, message=FALSE, warning=FALSE}
library(SamsaRaLight)
library(data.table)
```

### Set the stand geometry

You first need to precise the geometry of the stand with:

```{r stand_geom}
# Coordinates of the stand (for monthly radiation and rays geometry)
latitude <- 46 # The latitude of the stand (Y-coord in WGS85)
longitude <- 2 # The longitud eof the plot (X-coord in WGS85)

# The slope of the stand
slope <- 6

# Angle of slope bottom on the compass from the North, clockwise rotation (in degrees)
# northern aspect : 0, eastern aspect : 90, southern aspect : 180, western aspect : 270
aspect <- 144

# Angle from North to x axis clockwise. (in degrees)
# Default correspond to a Y axis oriented toward the North.
north_to_x_cw <- 54

# Considering a squared plot, number and size of the cells composing the grid
n_cells <-  10 # Number of cells within one length
cell_size <- 10 # Size of the length of a cell
```

### Create the trees dataset

To be able to run SamsaRaLight, you need to provide a well formatted
data.frame containing precise information about each tree of the plot.

The trees data.frame should contain all the below variables, with the
correct name and type of the column:

-   `id_tree`: Unique id of the tree. (integer)

-   `species`: Species Latin name. (character)

-   `x`, `y`, `z`: Coordinates of the base of the tree in the forest
    stand in m. (double)

-   `dbh_cm`: Diameter at breast height (1.30m) of the trunk of the tree
    in cm. (double)

-   `height_m`: Height of the tree trunk in m. (double)

-   `cbh_m`: Crown base height of the tree (i.e. height at wich the
    crown start) in m. (double)

-   `cradius_m`: Biggest radius of the tree crown in m (double)

-   `crown_type`: Type of the crown between "paraboloid" (1) and
    "ellispoid (2). (integer)

-   `crown_lad`: Leaf Area Density of the tree crown in m2/m3 (i.e.
    surface of leave per volume of crown, considering an homogeneous
    crown). Used when computing interception with a crown considered as
    a turbid medium. (double)

-   `crown_p`: Crown Openness of the tree (no unit) (i.e. Fraction of
    the energy of a light ray crossing the crown that is intercepted).
    Used when computing interception with a crown considered as a porous
    envelop. (double)

You can find an example tree dataset
`SamsaRaLight::data_trees_prenovel`. It is an uneven-aged mountain stand
of fir, spruce, beech located in Prenovel (Jura, France).

```{r data_trees}
trees <- as.data.table(SamsaRaLight::data_trees_prenovel) # Try to give a data.table to the function to avoid converting inside the function
trees
```

### Get monthly radiation

You will also need to provide a dataset with two radiation values for
each of the 12 months:

-   `month`: the month between 1 and 12

-   `Hrad`: monthly global radiation on a horizontal plane in MJ

-   `DGratio`: ratio of monthly diffuse energy to global energy (needed
    to compute direct energy)

You can fetch this dataset for a given latitude and longitude using the
function `get_monthly_rad()`. It fetches monthly data between given
start and end year (range between 2005 and 2020) from the PVGIS database
<https://joint-research-centre.ec.europa.eu/pvgis-photovoltaic-geographical-information-system_en>.
Here, you can choose to average the monthly values between all the years
between 2005 and 2020 or to choose values of a given year.

```{r monthly_rad}
monthly_rad <- SamsaRaLight::get_monthly_rad(latitude, longitude)
monthly_rad
```

### Run SamsaRaLight

Now, given the stand geometry, the trees dataset and the monthly
radiation, you can run SamsaRaLight using the function `sl_run()`.

```{r out_sl}
out <- SamsaRaLight::sl_run(
  trees, monthly_rad,
  latitude = latitude, slope = slope, 
  aspect = aspect, north_to_x_cw = north_to_x_cw,
  cell_size = cell_size, n_cells = n_cells,
  use_rcpp = T
)
```

The function returns a list with two dataframes:

`trees`: Light interception for each tree

-   `epot`: Potential energy intercepted by the tree without its
    neighbours (in MJ/year)

-   `e`: Energy intercepted by the tree when considering competition
    with neighbours (in MJ/year)

-   `lci`: Light competition index, computed as
    $lci = 1-\frac{e}{epot}$. It ranges from 0 to 1 with 0 being a tree
    without competition for light (all the energy that the tree can
    intercept from coming rays are intercepted) and 1 being a tree
    totally in competition for light (all the energy from the rays
    coming to the tree crown has been intercepted by the neighbours).

```{r}
summary(out$trees)
```

`cells`: Light coming to each cell of the plot

-   `e`: Total energy arriving to the cell (in MJ/year)

-   `erel`: Relative energy coming to the cell compared to the one above
    canopy. It ranges from 0 (no light on the ground) to 1 (no energy
    has been intercepted by above trees).

```{r}
summary(out$cells)
```

## Other parameters

Cf turbid medium, use_rcpp, use_torus...

## Speed comparison

### With 10m cells

```{r speed_10m}
microbenchmark::microbenchmark(
  "r" = SamsaRaLight::sl_run(
    trees, monthly_rad,
    latitude = latitude, slope = slope, 
    aspect = aspect, north_to_x_cw = north_to_x_cw,
    cell_size = 10, n_cells = 10,
    use_rcpp = FALSE
  ),
  "rcpp" = SamsaRaLight::sl_run(
    trees, monthly_rad,
    latitude = latitude, slope = slope, 
    aspect = aspect, north_to_x_cw = north_to_x_cw,
    cell_size = 10, n_cells = 10,
    use_rcpp = TRUE
  ),
  times = 10
)
```

### With 5m cells

```{r speed_5m}
# microbenchmark::microbenchmark(
#   "r" = SamsaRaLight::sl_run(
#     trees, monthly_rad,
#     latitude = latitude, slope = slope,
#     aspect = aspect, north_to_x_cw = north_to_x_cw,
#     cell_size = 5, n_cells = 20,
#     use_rcpp = FALSE
#   ),
#   "rcpp" = SamsaRaLight::sl_run(
#     trees, monthly_rad,
#     latitude = latitude, slope = slope,
#     aspect = aspect, north_to_x_cw = north_to_x_cw,
#     cell_size = 5, n_cells = 20,
#     use_rcpp = TRUE
#   ),
#   times = 10
# )
```
