---
title: "Non-axis-aligned rectangle stand from GPS data"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Non-axis-aligned rectangle stand from GPS data}
  %\VignetteEncoding{UTF-8}
  %\VignetteEngine{knitr::rmarkdown}
editor_options: 
  markdown: 
    wrap: 72
---

```{r include=FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r}
library(SamsaRaLight)
library(dplyr)
```

# Introduction

In previous tutorials, inventories were assumed to be defined directly in a local Cartesian coordinate
system (meters). In practice, however, forest inventories are often collected with GPS data. 

In this vignette, we how to convert an **inventory based on GPS coordinates** where tree positions are given as longitude and latitude. We will convert the geographic coordinates into planar and axis-align the inventory zone with the virtual plot axes.


# Context and data

We first use the example inventory **IRRES1**, stored in the package as
`SamsaRaLight::data_IRRES1`. This inventory was collected in Belgian
Ardennes by Gauthier Ligot in the scope of the IRRES project, which
investigates the transition from even-aged to uneven-aged forest
management. The stand is dense in a sloppy terrain and is mainly
composed of Norway spruce and Douglas-fir, with a coppice stool of beech
at its center and a few silver fir and larch trees.

```{r}
trees_irres <- SamsaRaLight::data_IRRES1$trees

str(trees_irres)
```

Running `check_inventory()` on this dataset immediately fails, with an
error indicating that columns `x` and `y` are missing. This is expected:
tree positions are provided as longitude (`lon`) and latitude (`lat`),
expressed in degrees. To illustrate why this is problematic, we
(incorrectly) rename longitude and latitude to `x` and `y` and attempt
to plot the inventory. It leads to a meaningless plot as angular
coordinates (degrees) are incompatible with crown dimensions expressed
in meters.

```{r}
SamsaRaLight::plot_inventory(
  trees_irres %>% rename(x = lon, y = lat)
)
```

# Convert the coordinates

Before creating a stand, coordinates must be converted into a planar
Cartesian system with metric units. Tree coordinates can be converted
from the global WGS84 reference system (EPSG:4326) to a projected
coordinate system expressed in meters such as the Universal Transverse
Mercator (UTM) system. It is a grid-based, metric coordinate system
mapping the Earth using 60 longitudinal zones (each of them being
6-degree between 80°S and 84°N latitude), for each of the two
hemispheres North and South.

The EPSG needed to convert coordinates depends on the plot coordinates.
However, the UTM zone can be automatically inferred from the mean
longitude ($zone = floor((lon + 180) / 6) + 1$ and hemisphere inferred
from the mean latitude ($hemisphere = 32600$ if latitude is positive or
$hemisphere = 32700$ if latitude is negative). Thus, EPSG code can be
automatically computed as $EPSG = hemisphere + zone$. The function
`SamsaRaLight::create_xy_from_lonlat()` allows to automatically convert
a data.frame containing lon/lat coordinates into planar XY coordinates
determining the appropriate UTM system.

```{r}
trees_irres_xy <- SamsaRaLight::create_xy_from_lonlat(trees_irres)

str(trees_irres_xy)
```

After this conversion, tree positions are expressed in meters and the
inventory can now be validated and visualized correctly.

```{r}
SamsaRaLight::check_inventory(trees_irres_xy$df)
plot_inventory(trees_irres_xy$df)
```


# Define the inventory zone

In the IRRES1 example dataset, the trees were inventoried within a
rectangular inventory zone. However, the vertices are also expressed in
a lon/lat coordinate system and therefore need to be converted.

```{r}
polygon_irres_xy <- SamsaRaLight::create_xy_from_lonlat(
  SamsaRaLight::data_IRRES1$core_polygon
)

str(polygon_irres_xy)
```

We can verify that the trees and the inventory polygon are both
expressed in metres within the same coordinate system. To do so, we can
use the same `plot_inventory()` function as above but adding the core
polygon data.frame as a second argument.

```{r}
SamsaRaLight::plot_inventory(
  trees_irres_xy$df,
  polygon_irres_xy$df
)
```

We can use the function `SamsaRaLight::check_polygon()` to check that
the core polygon is geometrically correct and encompasses all the
inventoried trees. If it does not, the function tries to correct it by
making minimal changes, such as converting the polygon into a valid one
(e.g. if the vertices are not in the correct order) or adding a small
buffer to the polygon in an attempt to include all the trees (e.g. if
some trees are close to the border, small rounding errors can lead to
the polygon excluding them computationally). Thus, the function returns
the minimally corrected polygon and specifies this with a message if the
polygon has been modified; otherwise, it throws an error.

```{r}
polygon_irres_xy$df <- SamsaRaLight::check_polygon(
  polygon_irres_xy$df,
  trees_irres_xy$df
)
```


# Create the virtual stand

Fortunately, the SamsaRaLight package allows you to provide both tree
inventory and core polygon tables with only longitude/latitude
coordinates to the `create_sl_stand()` function, which automatically
performs system conversions. The stand creation process will also handle
coordinate shifts into a relative coordinate system starting at 0.

Because the projected coordinates follow a conventional GIS orientation
(Y axis pointing North), we set `north2x = 90`, meaning that geographic
North corresponds to the positive Y direction.

```{r}
stand_irres <- SamsaRaLight::create_sl_stand(
  trees_inv = SamsaRaLight::data_IRRES1$trees,
  cell_size = 5,
  
  latitude = SamsaRaLight::data_IRRES1$info$latitude,
  slope    = SamsaRaLight::data_IRRES1$info$slope,
  aspect  = SamsaRaLight::data_IRRES1$info$aspect,
  north2x = 90,
  
  core_polygon_df = SamsaRaLight::data_IRRES1$core_polygon
)

plot(stand_irres)
```

The stand dimensions are chosen as the **smallest grid (in number of
cells)** that fully contains the inventory zone:

```{r}
stand_irres$geometry$n_cells_x
stand_irres$geometry$n_cells_y
```

This corresponds to a stand size of:

```{r}
stand_irres$geometry$n_cells_x * stand_irres$geometry$cell_size
stand_irres$geometry$n_cells_y * stand_irres$geometry$cell_size
```

Then, tree coordinates are shifted to a local coordinate system starting
at zero:

```{r}
stand_irres$transform$shift_x
stand_irres$transform$shift_y
```

# Set the axis-aligned rectangle option

At this stage, the inventory is not yet axis-aligned. This is not a
technical issue with this package, and the light computation can be run
using the virtual non-axis-aligned stand created above. However, as can
be seen in the above plot, the area surrounding the rectangular
inventory zone is empty, which could affect light interception.
Therefore, in most cases, it is preferable to work with a rectangle
aligned with the simulation axes. To do so, we have to recreate the
virtual stand by setting `modify_polygon = "aarect"` (for axis-aligned
rectangle), which:

1.  **compute the minimum bounding rectangle of the inventory polygon**
    (in this case, as our inventory zone is already a rectangle, it does
    not change anything),
2.  **rotate the entire stand (trees and polygon)** so that this
    rectangle becomes axis-aligned (the rotation counter-clockwise in
    degrees applied to the stand is stored internally in
    `transform$rotation_ccw$`)
3.  **update the `north2x` value accordingly** (and can be seen in the
    compass of the `plot()` function)

```{r}
stand_irres_aarect <- SamsaRaLight::create_sl_stand(
  trees_inv = SamsaRaLight::data_IRRES1$trees,
  cell_size = 5,
  
  latitude = SamsaRaLight::data_IRRES1$info$latitude,
  slope    = SamsaRaLight::data_IRRES1$info$slope,
  aspect  = SamsaRaLight::data_IRRES1$info$aspect,
  north2x = 90,
  
  core_polygon_df = SamsaRaLight::data_IRRES1$core_polygon,
  modify_polygon = "aarect"
)

plot(stand_irres_aarect)
```

As we can see, rotating the stand results in a rectangular inventory
zone (shown in yellow) that may not cover the entire virtual stand area.
This creates empty spaces around the borders of the virtual stand and
reduces the total basal area per hectare (due to a larger area with the
same number of trees). This could slightly bias the light computation,
even though the small empty areas on the borders could be negligible for
tree light interception. This can also be avoided by:

1.  Setting the cell size to a smaller value to reduce the empty space,
    which would result in much higher computation time.
2.  Alternatively, the rectangle could be set up without being
    axis-aligned and the 'fill_around' argument could be used. This will
    be explained in the second and third examples of this tutorial and
    involves filling around the inventory zone with virtual trees, but
    introduces stochasticity to the stand virtualisation.


# Run SamsaraLight

As shown in the previous tutorials, monthly radiation data are retrieved
using the geographic location of the stand.

```{r}
data_radiations_irres <- SamsaRaLight::get_monthly_radiations(
  latitude  = SamsaRaLight::data_IRRES1$info$latitude,
  longitude = SamsaRaLight::data_IRRES1$info$longitude
)
```

And the simulation is run using `run_sl()` (here run with the
axis-aligned inventory zone).

```{r}
output_irres_aarect <- SamsaRaLight::run_sl(
  sl_stand = stand_irres_aarect,
  monthly_radiations = data_radiations_irres
)

plot(output_irres_aarect)
```
