---
title: "1 - A first minimal case"
subtitle: "Compute tree light interception by symmetric crowns in an axis-aligned rectangle plot"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{1 - A first minimal case}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r echo=TRUE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```


This is the first tutorial, showing you how to solve a basic problem using the SamsaRaLight package. We will demonstrate this with the Prenovel dataset, which is stored in the package as `SamsaRaLight::data_prenovel`. This dataset represents an uneven-aged stand of fir, spruce and beech located in the Jura mountains in France. You can find more information in the data documentation.

We used this dataset as a minimal example this comes from a "marteloscope" exercise, that is a plot used for harvesting exercises on the field. The trees are inventoried inside a formalised 1-ha squared plot (100m*100m), thus being the easier inventory for using the SamsaRaLight package. In the next tutorials, we explain the case where trees are not inventoried in a formalized rectangle plot (4 - Create a virtual stand from more complex tree inventory).


```{r example, message=FALSE, warning=FALSE}
library(SamsaRaLight)
library(dplyr)
```



# Input data

## Format the tree inventory

The user needs to define the individual trees composing the stand with their crown dimensions. The data frame must be created with a specific format and variables. See documentation using `?SamsaRaLight::check_inventory()` to understand the format of the trees dataset, and validate it using the same function.

The dataframe must contain mandatory information about the trees: a unique integer id (`id_tree`), the species name (`species`), coordinates in meters (`x`, `y`) and diameter at breast height in cm (`dbh_cm`). The x and y coordinates must be given in meters on a flat plane (not considering slope), and the z coordinate will be computed automatically during the process from stand geometry. Be careful, you cannot use longitude/latitude (WGS84 coordinates system), as it is angular coordinates, thus not representing correctly relative distance between each tree. Consider converting the coordinates in a planar reference system such as local UTM. See the Tutorial 4 for an example on how to use SamsaRaLight from trees inventoried with GPS data giving long/lat coordinates (4 - Create a virtual stand from more complex tree inventory).

In this tutorial, we represent the tree crowns with simple shapes, defined in the `crown_type` column by a specific code: "E" (for an ellipsoid) or "P" (for a paraboloid). The ellipsoidal shape "E" is commonly used for broadleaved species, whereas a paraboloidal shape "P" is commonly used for conifers. When using simple crown shapes, the user must provide only 3 values to define the crown dimensions: the height of the tree (`h_m` in meters), the height of the base of the crown (`hbase_m` in meters) and the crown maximum radius that is the same is the four cardinal directions as we consider simple symmetric crowns (`rn_m`, `re_m`, `rs_m` and `rw_m`, all in meters). When considering those type of simple crown shapes ("E" or "P"), the user must not provide the `hmax_m` variable, which is the height at which the crown radius is maximum. Indeed, it is automatically computed during the process, being set to the crown base height ($hmax = hbase$) when considering a paraboloidal shape "P", and set at the middle of the crown for an ellipsoidal shape "E" ($hmax = h - 0.5*(h - hbase)$).  


```{r data_trees}
input_trees_inv <- SamsaRaLight::data_prenovel$trees
head(input_trees_inv)

# Check the format of the inventory table
SamsaRaLight::check_inventory(input_trees_inv)
```


The user can observe its validated inventory using the function `SamsaRaLight::plot_inventory()`. Computation time can be long as the trees are plotted in order of tree height to well represent the vertical structure of the stand.

```{r plot_inventory}
plot_inventory(input_trees_inv)
```

In the next tutorials, we will explain more in details the columns `crown_openness`/`crown_lad` (3 - Choose the transmission model) and consider more complex, asymmetric crown shapes (5 - Represent the crowns with more complex shapes).



## Create the SamsaRaLight input stand

The SamsaraLight ray-tracing model needs to be run on a axis-aligned rectangle stand that is split into square cells, with a given homogeneous slope, orientation and aspect. Moreover, the user needs to precise the plot latitude (Y-axis in WGS84 in degrees) as the model considers that light rays' geometry vary with plot latitude, discussed further in the Tutorial 2 (2 - Understand ray-tracing). 

The package provides a function that creates a standardized SamsaRaLight virtual stand as a `sl_stand` R S3 object: `SamsaRaLight::create_sl_stand()`. This function needs 5 mandatory arguments: the tree inventory table (`tree_inv`, formatted as described above), the size of the square cell that split the rectangle SamsaRaLight stand (`cell_size` in meters), the plot latitude (`latitude`) and three variables defining the plane stand geometry (`slope`, `aspect` and `north2x`, all in degrees). The function checks internally that all arguments are correct and well formatted (set the argument `verbose = FALSE` to avoid success messages).

Optionally but essential here to well define your virtual stand from the tree inventory, the user can define the inventory zone with the argument `core_polygon_df`. The polygon is defined by a data.frame that contains the relative coordinates `x` and `y` of the vertices of the zone where the trees have been inventoried.

The `cell_size` argument defines the precision of the SamsaRaLight ray-tracing model. The ray-tracing model will casts rays towards the center of each cell and estimate the light arriving to this cell after attenuation by tree crowns and the smaller the cells are defined, the more rays the model casts per m2. The default `cell_size` is 10, leading to very fast computation time. The user can go towards the maximum precision which is 1 (1m*1m square cells) to produce high-resolution shading maps (see the end of this tutorial), but will lead to much heavier computation time and memory use. Indeed, computation time is not proportional to `cell_size` but rather quadratic: the number of cells in the plot increases quadratically with the cell size, and also consequently is the total number of cast rays during the simulation (which is the main factor for computation time and memory use).    

The `north2x` argument is the angle from North to X-axis (clockwise), thus defining the rotation of real North (tree inventory) relative to the virtual plot X-axis (SamsaRaLight stand). The default 90° corresponds to a Y-axis oriented toward true North (0°: x-axis points North, 90° : x-axis points East, 180° : x-axis points South, 270° : x-axis points West). The `slope` and `aspect` arguments define the geometry in case of a non-plane stand, with the aspect variable representing the direction the slope faces downhill, defined by the angle of the steepest downslope direction measured clockwise from true North (0°: North-facing slope, East-facing slope: 90°, 180°: South-facing slope, 270°: West-facing slope). The Tutorial 2 (2 - Understand ray-tracing) also shows how the light on the ground varies with stand geometry. 


```{r stand_geom}
input_sl_stand <- SamsaRaLight::create_sl_stand(
  
  # Tree inventory formatted table
  trees_inv = input_trees_inv,
  
  # Precision of the SamsaRaLight model
  cell_size = 1,
  
  # Stand information
  latitude = SamsaRaLight::data_prenovel$info$latitude,
  slope = SamsaRaLight::data_prenovel$info$slope,
  aspect = SamsaRaLight::data_prenovel$info$aspect,
  north2x = SamsaRaLight::data_prenovel$info$north2x,
  
  # Definition of the inventory zone definition
  core_polygon_df = SamsaRaLight::data_prenovel$core_polygon
)
```


The user can observe the creating SamsaRaLight virtual stand object (`sl_stand` S3 object) using the base functions `print()`, `summary()` and `plot()`. The `plot()` function allows to observe the stand in two views: a stand map with circles representing the located maximum crown radius of trees from above, or a top-down plot mainly showing heights of crowns and trunks from a South-view (facing North) and a East-view (facing West).


```{r print_sl_stand}
print(input_sl_stand)
```

```{r summary_sl_stand}
summary(input_sl_stand)
```

```{r plot_sl_stand}
plot(input_sl_stand)
```

```{r plot_sl_stand_topdown}
plot(input_sl_stand, top_down = TRUE)
```



In the next tutorials, we explain how to create a virtual stand where trees are not inventoried directly in an axis-aligned rectangle plot (4 - Create a virtual stand from more complex tree inventory).



## Monthly radiations

The user needs to define in a data frame the monthly energy specific to its plot location. For each month (represented by an integer number between 1 and 12), one needs to inform $Hrad$ as the global monthly energies (in $MJ.m^{-2}$), and $DGratio$ as the ratio of diffuse energy relative to global energy (needed to represent the proportion of diffuse and direct energy). The discretization of diffuse and direct rays from the given radiation dataset is discussed in the next tutorial (2 - Understand ray-tracing).

This data frame can be automatically constructed using the SamsaRaLight function `SamsaRaLight::get_monthly_radiations()`, given the latitude and longitude of the plot. 
It gets radiation data from the PVGIS European database but needs an Internet connection. Otherwise, the monthly radiation data frame used in this tutorial (Prenovel stand) is stored within the package and can be retrieved using `SamsaRaLight::data_prenovel$radiations`.


```{r monthly_rad}
# Get the monthly radiation data frame
input_monthly_radiations <- SamsaRaLight::get_monthly_radiations(
  SamsaRaLight::data_prenovel$info$latitude, 
  SamsaRaLight::data_prenovel$info$longitude
)

input_monthly_radiations

# Check that it the format is correct
check_monthly_radiations(input_monthly_radiations)
```



# Run SamsaRaLight

From now, you can easily run the SamsaraLight ray-tracing model using the function `run_sl()` from the only two mandatory arguments: the created SamsaRaLight virtual stand object `sl_stand` and the monthly
radiations table `monthly_radiations`.

The `use_torus` argument is essential to account for the lack of information about trees surrounding the inventory zone. When set to TRUE (default), the stand is treated as an infinite plot, where the inventoried trees are surrounded by identical copies of the stand. When set to FALSE, the stand is assumed to be bordered by grasslands without trees, which may bias light interception estimates, particularly near the edges of the virtual stand. Thus, set the `use_torus` argument to FALSE only if it is biologically consistent, such as a stand in the middle of a clear-cut or surrounded by fields and roads.   

The package SamsaRaLight allows for running the simulation in parallel setting the argument `parallel_mode = TRUE`. When activating the parallel_mode, the simulation will take by default all the internal threads that are avaikable, but the user can decide how many cores to dedicate to the simulation with the argument `n_threads`. During the simulation, a message will appear to specify to the user how many cores are used by the simulation, and how many are available with the device, but the user can silence those messages with `verbose = TRUE` argument. The parallelisation is done internally across independent cells, while running sequentially each ray for a given cell. By doing so, the parallel mode is even essential when the number of cells is large. Parallelisation is done using OpenMP, but on platforms or compilers that do not support OpenMP, parallelisation is disabled and computations run in sequential mode (i.e. `parallel_mode = FALSE`).


```{r out_sl}
sl_output <- SamsaRaLight::run_sl(
    
    # Mandatory arguments
    sl_stand = input_sl_stand,
    monthly_radiations = input_monthly_radiations,
    
    # Surrounding environment
    use_torus = TRUE,
    
    # Parallelisation arguments
    parallel_mode = TRUE,
    n_threads = NULL
  )
```

For advanced users, an internal function is available using `SamsaRaLight:::run_sl_advanced()` allowing to control deeper ray-tracing parameters. The user-friendly function `SamsaRaLight::run_sl()` internally wrap the advenced version with default parameters. In the tutorials, we will not adress this advanced function, but all parameters with default values can be find following the documentation of this function. Otherwise, feel free to contact the main authors for a deeper use of the package and the model.



# Simulation output

The output is a complex S3 R object, with first a list of two elements: `$input` (that gathers inputs of the model defined above `$input$sl_stand` and `$input$monthly_radiations`, associated to the parameters of the simulation `$input$params`) and `$output` that essentially contains output light variables at the tree-level `output$light$trees` and at the cell-level `output$light$cells`. You can also observe `output$light$sensors` that are the light output at a located virtual light sensor, but this is discussed in the Tutorial 6 (6 - Add virtual sensors on the ground). Optionnaly, the user can set the `run_sl()` function argument `detailed_output` to TRUE in order to have more information about the rays discretisation and interceptions, but it is also discussed further in the next tutorial (2 - Understand ray-tracing). In this tutorial, we will focus here on both the tree- and cell-level output light variables, stored in `output$light`.  

The object `$output$cells` contains output light variables for each cell, identified by its unique id (`id_cell`, linked to the input cell dataframe `sl_output$input$sl_stand$cells` to retrive cell center coordinates). There are 3 output variables, which are `e` (for the energy arriving on the cell in MJ/m2), `pacl` (for proportion of above light canopy, which is the ratio between the energy arriving on the cell and the energy before interception by the trees) and `punobs` (for the proportion of energy on the cell that comes from unobstructed rays, i.e. rays that have not been intercepted by any trees).

```{r str_sl_output_cells}
str(sl_output$output$light$cells)
```

The object `$output$trees` contains output light variables for each trees, identified by its unique id (`id_tree`). There are 4 output variables, which are `e` (for the total energy intercepted by the tree in MJ), `epot` (for the potential energy intercepted by the tree without considering its neighbors in MJ, i.e. the total energy intercepted if the tree was alone with the same crown dimensions), $lci = 1 - e/epot$ (which is a light competition index and a good proxy for tree dynamics, see Beauchamp et al. 2025, representing the real intercepted energy compared to the potential energy it could intercept without competition) and `punobs` (for the proportion of energy intercepted by the tree that comes from unobstructed rays, i.e. rays that have not been intercepted by any other trees).

```{r str_sl_output_trees}
str(sl_output$output$light$trees)
```

The user can observe the output of the SamsaRaLight simulation object (`sl_output` S3 object) using the base functions `print()`, `summary()` and `plot()`. The `plot()` function allows to observe the stand map with tree- and cell-level colouring depending on the output light variables. By default, the plot shows the light competition index LCI for trees and the relative light on the ground PACL for cells, as there are good proxies for estimating tree and stand dynamics. In addition, the argument `show_trees` can be set to FALSE to plot only a shade map as a nice plot for representing the understorey light. Light variables to plot can be defined independently for both trees and cells by setting `what_trees` and `what_cells` arguments. For example, setting `what_trees = "intercepted"` and `what_cells = "absolute` allows to better catch the differences in light energies between trees and cells. However, be careful about the colours of the cells, white areas do no mean cells in high-light but brighter cells relatively to the others, it could be confusing without precise legend...  


```{r print_sl_output}
print(sl_output)
```

```{r summary_sl_output}
summary(sl_output)
```

```{r plot_sl_output}
plot(sl_output)
```

```{r plot_sl_output_notrees}
plot(sl_output, show_trees = F)
```

```{r plot_sl_output_aboslute}
plot(sl_output, what_trees = "intercepted", what_cells = "absolute")
```