---
title: "Tutorial 5 - Light sensors"
subtitle: "Calibrate species' leaf area density from light sensors"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Tuto5 - light sensors}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```


```{r example, message=FALSE, warning=FALSE}
library(SamsaRaLight)
library(dplyr)
library(ggplot2)
library(ggforce)
library(tidyr)
library(grid)
```


```{r}
# Set seed for randomization reproducibility
set.seed(5030)
```


# Inventory info

Test on IRRES1 inventory with all trees having a LAD of 0.5

```{r stand_geom}
# It is an uneven-aged mixed stand in Belgium with irregular crowns
data_trees <- SamsaRaLight::data_trees_irres1
data_rad <- SamsaRaLight::data_rad_irres1

# Coordinates of the stand (for monthly radiation and rays geometry)
latitude <- 50.04 # The latitude of the stand (Y-coord in WGS85)
longitude <- 5.2 # The longitude of the plot (X-coord in WGS85)

# The slope of the stand
slope <- 6.49

# Angle of slope bottom on the compass from the North, clockwise rotation (in degrees)
# northern aspect : 0, eastern aspect : 90, southern aspect : 180, western aspect : 270
aspect <- 43.54

# Angle from North to x axis clockwise. (in degrees)
# Default value 90 corresponds to a Y axis oriented toward the North.
north_to_x_cw <- 90
```


# Shift coordinates and define the rectangal square

```{r}
# Observe the plot coordinates
print("Range of the X- and Y-axis")
range(data_trees$x)
range(data_trees$y)

print("Width of the X- and Y-axis")
inv_size_x <- max(data_trees$x) - min(data_trees$x)
inv_size_x

inv_size_y <- max(data_trees$y) - min(data_trees$y)
inv_size_y

# Define the rectangle size
cell_size <- 10
n_cells_x <- 15
n_cells_y <- 11

plot_size_x <- cell_size * n_cells_x
plot_size_y <- cell_size * n_cells_y

# Shift coordinates by centering the plot
diff_x <- plot_size_x - inv_size_x
diff_y <- plot_size_y - inv_size_y

shift_x <- min(data_trees$x) - diff_x / 2 
shift_y <- min(data_trees$y) - diff_y / 2

data_trees$x <- data_trees$x - shift_x
data_trees$y <- data_trees$y - shift_y

print("New range of the X- and Y-axis")
range(data_trees$x)
range(data_trees$y)
```


# Fill around core polygon and run SamsaraLight

```{r out_samsa}
inv_filled <- SamsaRaLight::fill_inventory(data_trees, 
                                           cell_size, n_cells_x, n_cells_y)

out_filled <- SamsaRaLight::sl_run(
    inv_filled$new_inventory, data_rad,
    latitude = latitude, slope = slope, 
    aspect = aspect, north_to_x_cw = north_to_x_cw,
    cell_size = cell_size, 
    n_cells_x = n_cells_x, n_cells_y = n_cells_y,
    turbid_medium = TRUE,
    trunk_interception = FALSE
  )
```

```{r, fig.dim = c(8, 6)}
plot_inv_filled <- out_filled$cells %>% 
  dplyr::mutate(
    pacl_group = case_when(
      pacl_slope > 0.5 ~ "> 50%",
      pacl_slope > 0.25 ~ "25 - 50%",
      pacl_slope > 0.125 ~ "12.5 - 25%",
      pacl_slope > 0.0625 ~ "6.25 - 12.5%",
      pacl_slope <= 0.0625 ~ "< 6.25%"
    ),
    pacl_group = factor(pacl_group,
                        levels = c("< 6.25%", "6.25 - 12.5%",
                                   "12.5 - 25%", "25 - 50%",
                                   "> 50%"))) %>% 

ggplot() +
  
  geom_tile(aes(x = x_center, y = y_center,
                fill = pacl_group), color = "black") +
  scale_fill_grey(start = 0.2, end = 1,
                  drop = FALSE) +
  labs(fill = "Cells enlightment") +
  ggnewscale::new_scale_fill() +
  
  geom_ellipse(data = inv_filled$new_inventory, 
               mapping = aes(x0 = x, y0 = y, 
                             a = (re_m + rw_m) / 2,
                             b = (rn_m + rs_m) / 2,
                             angle = 0,
                             fill = species)) +
  coord_equal() +
  labs(fill = "Species") +
  scale_fill_manual(values = c("skyblue", "salmon", "yellow2", "olivedrab3", "mediumorchid1")) +
  
  scale_x_continuous(breaks = seq(0, n_cells_x*cell_size,
                                  by = cell_size),
                     labels = round(seq(0, n_cells_x*cell_size,
                                        by = cell_size),
                                    digits = 1)) +
  scale_y_continuous(breaks = seq(0, n_cells_y*cell_size,
                                  by = cell_size),
                     labels = round(seq(0, n_cells_y*cell_size,
                                        by = cell_size),
                                    digits = 1)) +
  
  geom_sf(data = inv_filled$core_polygon, inherit.aes = FALSE,
          fill = "yellow", color = "black", alpha = 0.7) +
  
  theme_minimal() +
  theme(panel.grid.minor = element_blank(),
        legend.position = "right") +
  xlab("") + ylab("")

plot_inv_filled
```

# Look for sensors and run SamsaraLight on them

```{r, fig.dim = c(8, 6)}
data_sensors <- data_sensors_irres1 %>% 
  dplyr::mutate(x = x - shift_x,
                y = y - shift_y)

plot_inv_filled_sensors <- plot_inv_filled +
  geom_rect(data = data_sensors,
            mapping = aes(xmin = x - cell_size/2,
                          ymin = y - cell_size/2,
                          xmax = x + cell_size/2,
                          ymax = y + cell_size/2),
            color = "black", fill = "gray", alpha = 0.5) +
  geom_text(data = data_sensors, 
            mapping = aes(x = x, y = y, label = id_sensor),
            size = 3)

plot_inv_filled_sensors
```

Set detailed output to TRUE because we need PACL in a horizontal plane, whereas the default PACL is in the slope.
Indeed, the sensor is not parallel to the slope, but horizontal.


```{r out_sensors}
out_sensors <- SamsaRaLight::sl_run(
    inv_filled$new_inventory, data_rad,
    sensors = data_sensors, sensors_only = TRUE,
    latitude = latitude, slope = slope, 
    aspect = aspect, north_to_x_cw = north_to_x_cw,
    cell_size = cell_size, 
    n_cells_x = n_cells_x, n_cells_y = n_cells_y,
    turbid_medium = TRUE,
    trunk_interception = FALSE,
    detailed_output = TRUE
  )
```


```{r time_sl}
microbenchmark::microbenchmark(
  "sl" = SamsaRaLight::sl_run(
    inv_filled$new_inventory, data_rad,
    sensors = data_sensors, sensors_only = FALSE,
    latitude = latitude, slope = slope, 
    aspect = aspect, north_to_x_cw = north_to_x_cw,
    cell_size = cell_size, 
    n_cells_x = n_cells_x, n_cells_y = n_cells_y,
    turbid_medium = TRUE,
    trunk_interception = FALSE
  ))
```

```{r time_sl_onlysensors}
microbenchmark::microbenchmark(
  "sl" = SamsaRaLight::sl_run(
    inv_filled$new_inventory, data_rad,
    sensors = data_sensors, sensors_only = TRUE,
    latitude = latitude, slope = slope, 
    aspect = aspect, north_to_x_cw = north_to_x_cw,
    cell_size = cell_size, 
    n_cells_x = n_cells_x, n_cells_y = n_cells_y,
    turbid_medium = TRUE,
    trunk_interception = FALSE
  ))
```


# Compare observed with predicted light from field and virtual sensors

```{r, fig.dim = c(8, 6)}
dplyr::left_join(
  data_sensors %>% 
    dplyr::select(id_sensor, pacl, pacl_direct, pacl_diffuse) %>% 
    dplyr::rename_at(vars(-"id_sensor"), ~paste0(., "_obs")),
  out_sensors$sensors %>% 
    dplyr::select(id_sensor, 
                  pacl = pacl_horizontal, 
                  pacl_direct = pacl_horizontal_direct, 
                  pacl_diffuse = pacl_horizontal_diffuse) %>% 
    dplyr::rename_at(vars(-"id_sensor"), ~paste0(., "_pred")),
  by = "id_sensor"
) %>% 
  dplyr::mutate(
    diff_pacl = pacl_obs - pacl_pred,
    diff_pacl_direct = pacl_direct_obs - pacl_direct_pred,
    diff_pacl_diffuse = pacl_diffuse_obs - pacl_diffuse_pred
  ) %>% 
  
  ggplot(aes(y = pacl_pred, x = pacl_obs)) +
  geom_point() +
  geom_smooth(method = "lm", formula = y ~ x, color = "salmon") +
  geom_abline(intercept = 0, slope = 1, color = "skyblue") +
  xlab("Observed PACL") +
  ylab("Predicted PACL with virtual sensors in SamsaraLight") +
  labs(title = "Comparison between PACL from field sensors and predicted with SamsaraLight",
       subtitle = "red line is a linear regression and blue line is the 1,1 control line for perfect fit") +
  theme_bw() +
  theme(plot.title = element_text(hjust = 0.5),
        plot.subtitle = element_text(hjust = 0.5))
```

# A first approach for LAD calibration

Above, the LAD parameter has been set to 0.5. Field sensors can be used to calibrate the tree leaf area density, an important parameter that controls the amount of light arriving on the ground.

To do so, a first approach could be to found a mean species' LAD value that minimize the mean residuals between observed and predicted total PACL.

```{r}
LADs <- seq(0.01, 2, by = 0.01)
```

```{r echo = T, results = 'hide'}
out_residuals <- vector("list", length(LADs))
i <- 0

for (lad in LADs) {
  
  print(lad)
  i <- i+1
  
  # Set the tested LAD value
  inv_test <- inv_filled$new_inventory %>% 
    dplyr::mutate(crown_lad = lad)
  
  # Run SamsaraLight
  out_sensors_test <- SamsaRaLight::sl_run(
    inv_test, data_rad,
    sensors = data_sensors, sensors_only = TRUE,
    latitude = latitude, slope = slope, 
    aspect = aspect, north_to_x_cw = north_to_x_cw,
    cell_size = cell_size, 
    n_cells_x = n_cells_x, n_cells_y = n_cells_y,
    turbid_medium = TRUE,
    trunk_interception = FALSE,
    detailed_output = TRUE
  )
  
  # Compute the mean residuals
  out_residuals[[i]] <- 
    
    dplyr::left_join(
      
      data_sensors %>% 
        dplyr::select(id_sensor, pacl, pacl_direct, pacl_diffuse) %>% 
        dplyr::rename_at(vars(-"id_sensor"), ~paste0(., "_obs")),
      
      out_sensors_test$sensors %>% 
        dplyr::select(id_sensor, 
                      pacl = pacl_horizontal, 
                      pacl_direct = pacl_horizontal_direct, 
                      pacl_diffuse = pacl_horizontal_diffuse) %>% 
        dplyr::rename_at(vars(-"id_sensor"), ~paste0(., "_pred")),
      
      by = "id_sensor"
    ) %>% 
    
    dplyr::mutate(
      res_pacl = pacl_obs - pacl_pred,
      res_pacl_direct = pacl_direct_obs - pacl_direct_pred,
      res_pacl_diffuse = pacl_diffuse_obs - pacl_diffuse_pred
    ) %>% 
    
    dplyr::summarise(across(c(res_pacl, res_pacl_direct, res_pacl_diffuse), mean)) %>% 
    dplyr::mutate(lad = lad) %>% 
    dplyr::select(lad, res_pacl, res_pacl_direct, res_pacl_diffuse)
}

out_residuals <- dplyr::bind_rows(out_residuals)
```

```{r, fig.dim = c(8, 6)}
best_lad <- out_residuals$lad[which.min(abs(out_residuals$res_pacl))]
text_best_lad <- textGrob(as.character(best_lad),
                          gp=gpar(fontsize=13, fontface="bold",
                                  col = "salmon"))  

ggplot(out_residuals, aes(y = res_pacl, x = lad)) +
  geom_point() +
  geom_hline(yintercept = 0, color = "darkgray") +
  geom_vline(xintercept = best_lad, color = "salmon") +
  annotation_custom(text_best_lad, 
                    xmin=best_lad, xmax=best_lad, 
                    ymin=-0.88, ymax=-0.88) + 
  coord_cartesian(clip = "off") +
  theme_bw() +
  ylab("Residuals between observed and predicted PACL") +
  xlab("Leaf area density LAD")
```


